---
output:
  pdf_document: default
  html_document: default
---
#Problem 2

#Calculate the Manhattan distance between any two vectors, vec1 and vec2

#Assign vectors, fill with any two vectors
#vec1 <- c(1,2,3,4)
#vec2 <- c(2,3,4,5)

Manhattan_dist <- function(vec1, vec2){
dist <- sum(abs(vec1 - vec2))
return (dist)
}

Manhattan_dist(vec1,vec2)

#Calculate the Euclidian distance between any two vectors, vec1 and vec2

#Assign vectors, fill with any two vectors
#vec1 <- c(1,2,3,4)
#vec2 <- c(2,3,4,5)

Euclidian_dist <- function(vec1, vec2){
dist <- sqrt(sum((vec1 - vec2)^2))
return (dist)
}

Euclidian_dist(vec1,vec2)
```{r}
#Problem #4

library(ggplot2)
data("mtcars")

#Calculate correlation between mpg and wt
cor_mpg_wt <- cor(mtcars$mpg, mtcars$wt)

#Plot correlation onto scatter plot
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  labs(title = "Correlation of MPG vs. Weight")
```
```{r}
#Problem #5

setwd("C:/Users/kathryn.poniatowski/OneDrive - Mass Transport Authority, Flint/UofM/CSC 587/Poniatowski-HW2")

#Import test data into dataframe
metabolite <- read.csv("metabolite.csv")

#Calculate missing precentage in columns
missing_percentage <- colMeans(is.na(metabolite))

#Drop columns with over 75% missing
metabolite_columns <- missing_percentage <= 0.75

#Assign columns to amended dataframe
metabolite_amended <- metabolite[, metabolite_columns]

# Replace missing values in remaining columns with the median of the column
metabolite_amended <- metabolite_amended
for (col in colnames(metabolite_amended)) {
  # Calculate the median excluding NA values
  median_value <- median(metabolite_amended[[col]], na.rm = TRUE)
  
  # Replace NA values with the median
  metabolite_amended[[col]][is.na(metabolite_amended[[col]])] <- median_value
}

head(metabolite_amended)
```
```{r}
library(ggplot2)
library(stats)

# Exclude non-numeric columns from the dataframe
numeric_data <- metabolite_amended[sapply(metabolite_amended, is.numeric)]

# Perform PCA on the numeric data (excluding the Label column if it exists)
pca <- prcomp(numeric_data, scale. = TRUE)

# Get the first two principal components (PC1, PC2)
pca_data <- as.data.frame(pca$x[, 1:2])

# Add the 'Label' column from the original dataframe to the PCA data
pca_data$Label <- metabolite_amended$Label

# Plotting the scatter plot of PC1 vs PC2, colored by 'Label'
ggplot(pca_data, aes(x = PC1, y = PC2, color = Label)) +
  geom_point() +
  labs(title = "Metabolite PCA Scatter Plot") +
  scale_color_manual(values = c("Group1" = "blue", "Group2" = "red"))

```


